A Simple Minibatch Acceptance Test for MCMC

### 2016-05-19 nips_2016.tex
Revised Appendics C and D. Only main thing I have in mind is to fix the conflicting temperature T definitions

Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} proposes an adaptive
minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
should be accepted or rejected based on the likelihood ratio. If the test cannot make a decision
over a certain confidence threshold, then they increase the minibatch size and test again. This
process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
the minibatch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties,
but comes at the cost of more computational time.

@@

\begin{itemize}[noitemsep]
    \item Some proof details on results in Section~\ref{sec:theory}.
    \item More information on Section~\ref{ssec:gaussians}.
    \item More information on Section~\ref{ssec:logistic}.
    \item More information on Section~\ref{ssec:nets}.
\end{itemize}

@@

{\color{blue}
Daniel: TODO
}

@@

{\color{blue}
Daniel: TODO
}

>>>>

Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} proposes an adaptive
minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
should be accepted or rejected based on the likelihood ratio. If the test cannot make a decision
over a certain confidence threshold, then they increase the minibatch size and test again. This
process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
the minibatch size; in the worst case, all $N$ data points may be needed in a single iteration.  A
similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties
based on concentration inequalities, but comes at the cost of more computational time.

@@

\begin{itemize}[noitemsep]
    \item Proofs for material in Section~\ref{sec:theory}.
    \item More information on Section~\ref{ssec:gaussians}.
    \item More information on Section~\ref{ssec:logistic}.
    \item More information on Section~\ref{ssec:nets}.
\end{itemize}

@@

In this section, we go over some details of our logistic regression experiment. The feature vector
for an image consists of its pixel values, normalized between 0 and 1. For simplicity, we only
consider the binary classification case, so we only use digits one (denoted as output $y = 1$) and
seven (denoted as $y = -1$). The probability of the $i^{\rm th}$ output $y_i \in \{-1,+1\}$ with the
input vector $x_i$ is
\begin{equation}
p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
\end{equation}
where $\theta$ is the 784-length parameter vector.

For our experiment, we impose a uniform prior to represent our lack of knowledge about $\theta$.  We
use a random walk proposer, which can be modeled as $\theta' = \theta_i + \mathcal{N}(0,
\sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
the variance to be a constant $\sigma^2 = 0.01$ for all components. We initialize $\theta_0$ to be a
vector of all ones, and set our minibatch size as $m=50$.

For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we
initialize the temperature $T_0 = 3000$, and decrease it by the rule $T_i =  T_0 /(i + 1)^{3} \mbox{
for } i \leq N/2 \mbox{ and } = T_0 /(i + 1)^{0.2} \mbox{ for } i >N/2 $, where $N=4000$ is
the total number of iterations. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
current iteration. Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
versus iteration count.

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{./figures/appendixExp2}
  \caption{Additional results for the logistic regression experiment.}
  \label{fig:appendixExp2}
\end{figure}

For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
the actual minibatch size at the end of each iteration in adaptive MH testing.

@@

For our neural network experiment, we use the architecture discussed in Section~\ref{ssec:nets}. We
use the SGHMC as the proposer for our minibatch MH testing. For the baseline, we use a tuned
adaptive gradient descent optimizer, whose step size changes by $0.01 /(i+1)^{0.4}$. Both methods
have minibatch sizes set at $m=200$. There are one million total data instances $x_i$.

For SGHMC, we use the simplified update equations~\cite{sghmc_2014}:
\begin{equation}
\Delta \theta = v, \quad \quad\Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha
-\hat{\beta}) \eta),
\end{equation}
where $v$ represents auxiliary momentum variables, and $\nabla U(\theta)$ is the gradient of the
system. We set the hyperparameters to be $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$. We use the
empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$.

In order to control ${\rm std}(\Delta')$, we initialize the temperature at 1000, and adjust it
at iteration $i$ according to $T_i = \max\{1, 1000/(i+1)^{0.5}\}$.
